Bayesian exponential random graph modeling of KWL Burial network

### 2020-11-13 analysis/paper/paper.Rmd
add a reference for MCMC

Our pericolonial impact hypothesis is that restricted grave goods or accumulation of goods present in a few individuals after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access high value goods. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to the interest of burial goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power. If the social inequality gradually increased in Kiwulan, then we would observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status). Network analysis provides a way to examine the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.

@@

<!--Network construction and specification-->
We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language with the bergm package [@Caimo2014], which built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008]. Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. Based on our hypothesis, we model a network with increased social inequality that represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects. We modeled several node attributes for homophily effects, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall. We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried close to each other. Our model infers the presence of a few individuals as centers having more relations with others, which hints the emergence of social inequality. We expect that observed network after the European arrival will have a better match with our model compared with the network before European contact.

<!--Model fitting and prior setting--> 
After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017].

>>>>

Our pericolonial impact hypothesis is that restricted grave goods or accumulation of goods present in a few individuals after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access high value goods. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to the interest of burial goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power. If the social inequality gradually increased in Kiwulan, then we would observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status). ERGM in a Bayesian framework provides a way to examine the formation of network ties and its underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.

@@

<!--Network construction and specification-->
We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language with the bergm package [@Caimo2014], which built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008]. Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. Based on our hypothesis, we model a network with increased social inequality that represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects. We modeled four node attributes for homophily effects, including age, gender, ritual activity, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall. We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@@LiandWu2006]. Our model infers the presence of a few individuals as centers having more relations with others, which hints the emergence of social inequality. We expect that observed network after the European arrival will have a better match with our model compared with the network before European contact.

<!--Model fitting and prior setting--> 
After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution[@Hamra2013]. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017].
